---
title: 'Practical Machine Learning Assignment: Predicting Class of Excercise'
author: "Tom Zalenski"
date: "March 23, 2019"
output:
  html_document: null
  word_document: default
  keep_md: yes
  pdf_document: default
  fig_width: 9
---
    
# Synopsis:

The goal of this report is to use raw data from sensors worn by the user and attached to the equipment, to predict the exercise class [A,B,C,D,E].
The sensors are:

    1.  roll, pitch & yaw of users belt
    2.  X, Y & Z axis from gyro on users belt
    3.  X, Y & Z axis from accelerometer on users belt     
    4.  X, Y & Z axis from magnets on users belt
    5.  roll, pitch & yaw of users arm
    6.  X, Y & Z axis from gyro on users arm
    7.  X, Y & Z axis from accelerometer on users arm
    8.  X, Y & Z axis from magnets on users arm
    9.  roll, pitch & yaw of dumbbell
    10. X, Y & Z axis from gyro on dumbbell
    11. X, Y & Z axis from accelerometer on dumbbell
    12. X, Y & Z axis from magnets on dumbbell
    13. roll, pitch & yaw of users forearm
    14. X, Y & Z axis from gyro on users forearm
    15. X, Y & Z axis from accelerometer on users forearm
    16. X, Y & Z axis from magnets on users forearm


## Loading the source data:

```{r, message=FALSE, results=FALSE, cache=TRUE }
library(dplyr)
library(caret)

train_df <- read.csv( "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv" )
test_df  <- read.csv( "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" )
```

## Preprocessing the data:

Note, the train_df data has the outcome variable "classe", while the test_df does not.
For the purpose of model building and testing, a subset of data containing the outcome, must be created. 
The following code subdivides the original training data into two subsets:
 
```{r, cache=TRUE }
set.seed(3433)
inTrain  = createDataPartition( train_df$classe, p = 3/4)[[1]]
training = train_df[ inTrain,]
testing  = train_df[-inTrain,]
```

This section of code creates a formula with "classe" as the outcome and all raw sensor data as predictors:

```{r, cache=TRUE }
raw_sensor_names <- c("roll_belt",         "pitch_belt",        "yaw_belt",          
                      "gyros_belt_x",      "gyros_belt_y",      "gyros_belt_z",        
                      "accel_belt_x",      "accel_belt_y",      "accel_belt_z",     
                      "magnet_belt_x",     "magnet_belt_y",     "magnet_belt_z",      
                      "roll_arm",          "pitch_arm",         "yaw_arm",                           
                      "gyros_arm_x",       "gyros_arm_y",       "gyros_arm_z",       
                      "accel_arm_x",       "accel_arm_y",       "accel_arm_z",        
                      "magnet_arm_x",      "magnet_arm_y",      "magnet_arm_z",                 
                      "roll_dumbbell",     "pitch_dumbbell",    "yaw_dumbbell",        
                      "gyros_dumbbell_x",  "gyros_dumbbell_y",  "gyros_dumbbell_z",  
                      "accel_dumbbell_x",  "accel_dumbbell_y",  "accel_dumbbell_z", 
                      "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", 
                      "roll_forearm",      "pitch_forearm",     "yaw_forearm",                 
                      "gyros_forearm_x",   "gyros_forearm_y",   "gyros_forearm_z",   
                      "accel_forearm_x",   "accel_forearm_y",   "accel_forearm_z",           
                      "magnet_forearm_x",  "magnet_forearm_y",  "magnet_forearm_z" )

raw_sensor_variables <- paste( raw_sensor_names, collapse="+")
f <- paste("classe ~", raw_sensor_variables )
```

Now multiple classification models can be tried:

```{r, results=FALSE, cache=TRUE }
mod_rf  <- train( as.formula(f), data = training, method = "rf"  )
mod_gbm <- train( as.formula(f), data = training, method = "gbm" )
mod_lda <- train( as.formula(f), data = training, method = "lda" )

test_pred_rf  <- predict( mod_rf,  testing )
test_pred_gbm <- predict( mod_gbm, testing )
test_pred_lda <- predict( mod_lda, testing )

test_predDF <- data.frame( pred_rf  = test_pred_rf,
                           pred_gbm = test_pred_gbm,
                           pred_lda = test_pred_lda, class = testing$classe )
test_combModFit <- train( class ~ ., method = "rf", data = test_predDF )
test_combPred   <- predict( test_combModFit, test_predDF )
```

## Results and Summary:

```{r, cache=TRUE }
confusionMatrix( test_combPred, testing$classe )
```

The overall accuracy using the training data exceeds 99%. Now to use evaluate the actual testing data and predict the excerise class:

```{r, cache=TRUE }
final_test_pred_rf  <- predict( mod_rf,  test_df )
final_test_pred_gbm <- predict( mod_gbm, test_df )
final_test_pred_lda <- predict( mod_lda, test_df )
final_test_predDF   <- data.frame( pred_rf  = final_test_pred_rf,
                                   pred_gbm = final_test_pred_gbm, 
                                   pred_lda = final_test_pred_lda )
final_test_combPred <- predict( test_combModFit, final_test_predDF )
final_test_combPred
```

